### 2020-06-25 submission/response_to_reviewers.md

**Reviewer #1**

**The authors put together an interesting 10 simple rules article sharing their experiences with code clubs.**

**I find the proposed code clubs an interesting idea and would be interested in trying this out in the lab. However, quite a bit more practical advice from this manuscript would be helpful to get up and running. Practical advice can be in the form of things that worked, but also explicitly mentioning what didn't work so that others can learn from the authors experience.**

> Thank you for the feedback on where the manuscript could provide more context. Throughout their review, the reviewer encourages us to make more comments about groups where there are more people with CS backgrounds. Admittedly, our lab is mostly made of researchers with non-CS backgrounds and so our implementation of Code Clubs is with that audience in mind. We have added text as appropriate to respond to the following suggestions. We have expanded our listing of example code clubs in Table 1 along with links to GitHub repositories where they were used. Examples of Code Clubs that did not work are highlighted throughout the manuscript (e.g. teaching Julia or not clearly stating goals).

### 2020-06-25 submission/manuscript.Rmd
Incorporate edits

The standard undergraduate and graduate training in the biological sciences is typically insufficient for developing these skills [@Barone2017]. To fill this void, there has been growth in online data science training resources through companies including Data Camp (https://www.datacamp.com) and Codecademy (https://www.codecademy.com), massive open online courses (MOOCs), and in person workshops hosted by organizations including The Carpentries (https://carpentries.org). Each of these formats are highly popular among their audience (e.g. [@Wilson2016]); however, recent analyses of workshops suggests that they have little impact on the long term retention of the material that is taught [@Feldon2017]. This result should not be surprising. These learning formats ask participants to engage in massed practice; an approach that is known to be ineffective [@Carpenter2012; @Rohrer2015b; @Bud2010; @Karpicke2011]. In contrast, participants need regular opportunities to engage in repeated deliberate practice that encourage them to practice their retrieval and application skills of the material [@Rohrer2015a; @Karpicke2008]. There is clearly a need to develop a framework that can build upon material introduced in coursework and workshops and that can help researchers to stay up to date on the latest best practices, algorithms, and tools.

To address this need, we noted the similarity between the feeling of unease in analyzing the data and the struggles scientists face with engaging the voluminous scientific literature. A common strategy for keeping up with the literature is by participating in a journal club. Regardless of the discipline, a Journal Club involves group discussion of a pre-selected paper that range from informal discussions to PowerPoint presentations and course credit. In addition to building upon material from traditional coursework and staying current on the literature, journal clubs help strengthen skills in critical thinking, communication, and integrating the literature [@Lonsdale2016]. Since most journal clubs occur on a regular schedule, they are effective by virtue of repeated practice. With this model in mind, over the past four years we have experimented with creating a Code Club model with the goal of improving reproducible data analysis skills in a community environment.

@@

From these experiences, we began a collective conversation to improve our Code Club model. We have identified two successful approaches. The first is a more constructive version of a group code review. The presenter clearly states the problem they want to solve, breaks the participants into smaller groups, and then asks each group to solve the problem, or a portion of it. For example, someone may have an R script with a repeating chunk of code. The challenge for the session would be to convert the code chunk into a function to be called throughout the script to make it "DRY" (i.e. Don't Repeat Yourself [@Wilson2014]). The presenter leaves the session with several partial or working solutions to their problem and the importance of writing DRY code is reinforced. The second approach is a tutorial. The presenter introduces a new package or technique and assigns an activity to practice the new approach. For example, at one Code Club participants were given raw data and a finished plot. Paired participants were tasked with generating the plot from the data using R syntax from either the base language or the ggplot2 package. For this exercise, base R users had to use ggplot2 and vice versa. In either approach, the Code Club ends with a report back to the larger group describing the approach each pair took. We generally find that preparing a Code Club session takes similar effort to preparing a journal club presentation. Our Code Clubs typically have 7 to 10 participants, but the inherent "think-pair-share" approach should allow it to be scaled to larger groups [@Kothiyal2014].

We continue to experiment with approaches for running Code Club. For example, during the early period of the Covid-19 pandemic in 2020 when many research labs were shuttered, PDS started posting Code Club materials (https://www.riffomonas.org/code_club/) that individuals or research groups can use to practice their coding skills. During this time, we continued our group's Code Club sessions using the breakout room and screen sharing features in Zoom. Regardless of our approach to leading a Code Club, we have learned that it is critical for the presenter to clearly articulate their goals and facilitate participant engagement. Although some Code Club sessions may be more experimental than others, on the whole they are a critical tool to train bench scientists in reproducible data analysis practices. We have provided some examples of successful topics in Table 1. We have summarized the results of our own experimentation as Code Club presenters and participants into Ten Rules. The first 3 rules apply to all participants, Rules 4 through 8 are targeted to presenters, and the last two for non-presenters. As the rules describe, this model can easily be adapted to groups of non-biologists or those with stronger backgrounds in computer science.

@@

As suggested by Rule 7, a significant challenge to presenting at Code Club is selecting topics and activities that appeal to a critical mass of the participants. This is particularly difficult if participants have a wide range of coding experience, which can follow a turnover in group membership. Beginners will benefit from sessions that cover fundamental concepts and functions. The benefits for more experienced participants include the improved understanding of concepts by teaching and breaking down problems into simpler elements. Furthermore, instead of focusing on core functions, the group could balance Code Club sessions that cover basics with those that introduce new methods and packages to the group. We have also identified several strategies to overcome the challenges presented by participants at varying skill levels. Central to the Code Club format is the use of paired-programming [@Hanks2011]. Instead of letting participants form their own pairs, the presenter can select pairs of participants with either similar or differing skill levels, depending on the presenter's goals. Partnerships between those with similar skill levels requires the presenter to design appropriate activities for each skill level. We have found that commenting code is a good skill for beginners since it forces them to dissect and understand a code chunk line-by-line. It also reinforces the value of commenting as they develop their skills and independence. An advantage of forming partnership between people with disparate skill levels is that it is more likely for groups to provide the presenter with a diverse range of methods that achieve the same result. This approach to pairing also helps to graft new members that have emerging programming skills. Regardless of how partners are selected, consider asking the pairs to identify a navigator and a driver [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type, thus ensuring participation of both partners. Midway through the activity, the presenter can have the partners switch roles. Intentionally forming pairs can also engineer group interactions by avoiding potentially disruptive partnerships or pairing reliable role models with new group members.

@@

The most important rules are the first and last. Members of the Code Club need to feel comfortable with other group members and sufficiently empowered to try something new or ask for help. Aside from expanding our programming skills, we have noticed two other benefits that help create a positive culture. First, we have intentionally interviewed postdoc candidates on the days we hold Code Club. We make it clear that they are not being assessed on their coding skills, but instead use it as an opportunity to see how the candidate interacts with other members of the research group. At the same time, the candidate can learn about the culture of the research group through active participation. Second, members of other research groups have integrated themselves into our Code Club to minimize the isolation they feel in growing their skills within smaller research groups. This speaks to both the broader need for Code Club and the likelihood of success when expanded to include a larger group of individuals with broader research interests. There is no reason that the Code Club format would not work with groups independent of a research group as long as they follow the Rules. Finding datasets and applications that are interesting to a critical mass of people will likely be key to starting and sustaining such a group. Ultimately, Code Club has improved the overall data analysis skills, community, and research success of our lab by empowering researchers to seek help from their colleagues.

>>>>

The standard undergraduate and graduate training in the biological sciences is typically insufficient for developing these skills [@Barone2017]. To fill this void, there has been growth in online data science training resources through companies including Data Camp (https://www.datacamp.com) and Codecademy (https://www.codecademy.com), massive open online courses (MOOCs), and in person workshops hosted by organizations including The Carpentries (https://carpentries.org). Each of these formats are highly popular among their audience (e.g. [@Wilson2016]); however, recent analyses of workshops suggests that they have little impact on the long term retention of the material that is taught [@Feldon2017]. This result should not be surprising. These learning formats ask participants to engage in massed practice; an approach that is known to be ineffective [@Carpenter2012; @Rohrer2015b; @Bud2010; @Karpicke2011]. In contrast, participants need regular opportunities to engage in repeated deliberate practice that encourage them to practice their retrieval and application of the material [@Rohrer2015a; @Karpicke2008]. There is clearly a need to develop a framework that can build upon material introduced in coursework and workshops and that can help researchers stay up to date on the latest best practices, algorithms, and tools.

To address this need, we noted the similarity between the feeling of unease in analyzing the data and the struggles scientists face with engaging the voluminous scientific literature. A common strategy for keeping up with the literature is to participate in a Journal Club. Regardless of the discipline, a Journal Club involves group discussion of a pre-selected paper that range from informal discussions to PowerPoint presentations and course credit. In addition to building upon material from traditional coursework and staying current on the literature, Journal Clubs help strengthen skills in critical thinking, communication, and integrating the literature [@Lonsdale2016]. Since most Journal Clubs occur on a regular schedule, they are effective by virtue of repeated practice. With this model in mind, over the past four years we have experimented with creating a Code Club model with the goal of improving reproducible data analysis skills in a laboratory environment.

@@

From these experiences, we began a collective conversation to improve our Code Club model. We have identified two successful approaches. The first is a more constructive version of a group code review. The presenter clearly states the problem they want to solve, breaks the participants into smaller groups, and then asks each group to solve the problem, or a portion of it. For example, someone may have an R script with a repeating chunk of code. The challenge for the session would be to convert the code chunk into a function to be called throughout the script to make it "DRY" (i.e. Don't Repeat Yourself [@Wilson2014]). The presenter leaves the session with several partial or working solutions to their problem and the importance of writing DRY code is reinforced. The second approach is a tutorial. The presenter introduces a new package or technique and assigns an activity to practice the new approach. For example, at one Code Club participants were given raw data and a finished plot. Paired participants were tasked with generating the plot from the data using R syntax from either the base language or the ggplot2 package. For this exercise, base R users had to use ggplot2 and vice versa. In either approach, the Code Club ends with a report back to the larger group describing the approach each pair took. We generally find that preparing a Code Club session takes similar effort to preparing a Journal Club presentation. Our Code Clubs typically have 7 to 10 participants, but the inherent "think-pair-share" approach should allow it to be scaled to groups of variable sizes [@Kothiyal2014].

We continue to experiment with approaches for running Code Club. For example, during the COVID-19 pandemic in 2020 when many research labs were shuttered, PDS started posting Code Club materials (https://www.riffomonas.org/code_club/) that individuals or research groups can use to practice their coding skills. During this time, we continued our group's Code Club sessions using the breakout room and screen sharing features in Zoom. Regardless of our approach to leading a Code Club, we have learned that it is critical for the presenter to clearly articulate their goals and facilitate participant engagement. Although some Code Club sessions may be more experimental than others, on the whole they are a critical tool to train bench scientists in reproducible data analysis practices. We have provided some examples of successful topics in Table 1. We have summarized the results of our own experimentation as Code Club presenters and participants into Ten Rules. The first 3 rules apply to all participants, Rules 4 through 8 are targeted to presenters, and the last two for non-presenters. As the rules describe, this model can easily be adapted to groups of non-biologists or those with stronger backgrounds in computer science.

@@

As suggested by Rule 7, a significant challenge to presenting at Code Club is selecting topics and activities that appeal to a critical mass of the participants. This is particularly difficult if participants have a wide range of coding experience, which can follow a turnover in group membership. Beginners will benefit from sessions that cover fundamental concepts and functions. The benefits for more experienced participants include the improved understanding of concepts by teaching and breaking down problems into simpler elements. Furthermore, they can benefit by contributing to an environment of learning that they previously benefited from. In addition, instead of focusing on core functions, the group could balance Code Club sessions that cover basics with those that introduce new methods and packages to the group. We have also identified several strategies to overcome the challenges presented by participants at varying skill levels. Central to the Code Club format is the use of paired-programming [@Hanks2011]. Instead of letting participants form their own pairs, the presenter can select pairs of participants with either similar or differing skill levels, depending on the presenter's goals. Partnerships between those with similar skill levels requires the presenter to design appropriate activities for each skill level. We have found that commenting code is a good skill for beginners since it forces them to dissect and understand a code chunk line-by-line. It also reinforces the value of commenting as they develop their skills and independence. An advantage of forming partnership between people with disparate skill levels is that it is more likely for groups to provide the presenter with a diverse range of methods that achieve the same result. This approach to pairing also helps to graft new members that have emerging programming skills. Regardless of how partners are selected, consider asking the pairs to identify a navigator and a driver [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type, thus ensuring participation of both partners. Midway through the activity, the presenter can have the partners switch roles. Intentionally forming pairs can also engineer group interactions by avoiding potentially disruptive partnerships or pairing reliable role models with new group members.

@@

The most important rules are the first and last. Members of the Code Club need to feel comfortable with other group members and sufficiently empowered to try something new or ask for help. Aside from expanding our programming skills, we have noticed two other benefits that help create a positive culture. First, we have intentionally interviewed postdoctoral candidates on the days we hold Code Club. We make it clear that they are not being assessed on their coding skills, but instead use it as an opportunity to see how the candidate interacts with other members of the research group. At the same time, the candidate can learn about the culture of the research group through active participation. Second, members of other research groups have integrated themselves into our Code Club to minimize the isolation they feel in growing their skills within smaller research groups. This speaks to both the broader need for Code Club and the likelihood of success when expanded to include a larger group of individuals with broader research interests. There is no reason that the Code Club format would not work with groups independent of a research group as long as everyone follows the Rules. Finding datasets and applications that are interesting to a critical mass of people is essential to starting and sustaining such a group. Ultimately, Code Club has improved the overall data analysis skills, community, and research success of our lab by empowering researchers to seek help from their colleagues.

### 2020-06-25 submission/manuscript.tex
Incorporate edits

The standard undergraduate and graduate training in the biological
sciences is typically insufficient for developing these skills {[}2{]}.
To fill this void, there has been growth in online data science training
resources through companies including Data Camp
(\url{https://www.datacamp.com}) and Codecademy
(\url{https://www.codecademy.com}), massive open online courses (MOOCs),
and in person workshops hosted by organizations including The
Carpentries (\url{https://carpentries.org}). Each of these formats are
highly popular among their audience (e.g.~{[}3{]}); however, recent
analyses of workshops suggests that they have little impact on the long
term retention of the material that is taught {[}4{]}. This result
should not be surprising. These learning formats ask participants to
engage in massed practice; an approach that is known to be ineffective
{[}5--8{]}. In contrast, participants need regular opportunities to
engage in repeated deliberate practice that encourage them to practice
their retrieval and application skills of the material {[}9,10{]}. There
is clearly a need to develop a framework that can build upon material
introduced in coursework and workshops and that can help researchers to
stay up to date on the latest best practices, algorithms, and tools.

To address this need, we noted the similarity between the feeling of
unease in analyzing the data and the struggles scientists face with
engaging the voluminous scientific literature. A common strategy for
keeping up with the literature is by participating in a journal club.
Regardless of the discipline, a Journal Club involves group discussion
of a pre-selected paper that range from informal discussions to
PowerPoint presentations and course credit. In addition to building upon
material from traditional coursework and staying current on the
literature, journal clubs help strengthen skills in critical thinking,
communication, and integrating the literature {[}11{]}. Since most
journal clubs occur on a regular schedule, they are effective by virtue
of repeated practice. With this model in mind, over the past four years
we have experimented with creating a Code Club model with the goal of
improving reproducible data analysis skills in a community environment.

@@

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}12{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using R syntax from either the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. We generally
find that preparing a Code Club session takes similar effort to
preparing a journal club presentation. Our Code Clubs typically have 7
to 10 participants, but the inherent ``think-pair-share'' approach
should allow it to be scaled to larger groups {[}13{]}.

We continue to experiment with approaches for running Code Club. For
example, during the early period of the Covid-19 pandemic in 2020 when
many research labs were shuttered, PDS started posting Code Club
materials (\url{https://www.riffomonas.org/code_club/}) that individuals
or research groups can use to practice their coding skills. During this
time, we continued our group's Code Club sessions using the breakout
room and screen sharing features in Zoom. Regardless of our approach to
leading a Code Club, we have learned that it is critical for the
presenter to clearly articulate their goals and facilitate participant
engagement. Although some Code Club sessions may be more experimental
than others, on the whole they are a critical tool to train bench
scientists in reproducible data analysis practices. We have provided
some examples of successful topics in Table 1. We have summarized the
results of our own experimentation as Code Club presenters and
participants into Ten Rules. The first 3 rules apply to all
participants, Rules 4 through 8 are targeted to presenters, and the last
two for non-presenters. As the rules describe, this model can easily be
adapted to groups of non-biologists or those with stronger backgrounds
in computer science.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. Beginners will benefit from sessions that cover fundamental
concepts and functions. The benefits for more experienced participants
include the improved understanding of concepts by teaching and breaking
down problems into simpler elements. Furthermore, instead of focusing on
core functions, the group could balance Code Club sessions that cover
basics with those that introduce new methods and packages to the group.
We have also identified several strategies to overcome the challenges
presented by participants at varying skill levels. Central to the Code
Club format is the use of paired-programming {[}15{]}. Instead of
letting participants form their own pairs, the presenter can select
pairs of participants with either similar or differing skill levels,
depending on the presenter's goals. Partnerships between those with
similar skill levels requires the presenter to design appropriate
activities for each skill level. We have found that commenting code is a
good skill for beginners since it forces them to dissect and understand
a code chunk line-by-line. It also reinforces the value of commenting as
they develop their skills and independence. An advantage of forming
partnership between people with disparate skill levels is that it is
more likely for groups to provide the presenter with a diverse range of
methods that achieve the same result. This approach to pairing also
helps to graft new members that have emerging programming skills.
Regardless of how partners are selected, consider asking the pairs to
identify a navigator and a driver {[}16{]}. The driver types at the
computer while the navigator tells them what to type, thus ensuring
participation of both partners. Midway through the activity, the
presenter can have the partners switch roles. Intentionally forming
pairs can also engineer group interactions by avoiding potentially
disruptive partnerships or pairing reliable role models with new group
members.

>>>>

The standard undergraduate and graduate training in the biological
sciences is typically insufficient for developing these skills {[}2{]}.
To fill this void, there has been growth in online data science training
resources through companies including Data Camp
(\url{https://www.datacamp.com}) and Codecademy
(\url{https://www.codecademy.com}), massive open online courses (MOOCs),
and in person workshops hosted by organizations including The
Carpentries (\url{https://carpentries.org}). Each of these formats are
highly popular among their audience (e.g.~{[}3{]}); however, recent
analyses of workshops suggests that they have little impact on the long
term retention of the material that is taught {[}4{]}. This result
should not be surprising. These learning formats ask participants to
engage in massed practice; an approach that is known to be ineffective
{[}5--8{]}. In contrast, participants need regular opportunities to
engage in repeated deliberate practice that encourage them to practice
their retrieval and application of the material {[}9,10{]}. There is
clearly a need to develop a framework that can build upon material
introduced in coursework and workshops and that can help researchers
stay up to date on the latest best practices, algorithms, and tools.

To address this need, we noted the similarity between the feeling of
unease in analyzing the data and the struggles scientists face with
engaging the voluminous scientific literature. A common strategy for
keeping up with the literature is to participate in a Journal Club.
Regardless of the discipline, a Journal Club involves group discussion
of a pre-selected paper that range from informal discussions to
PowerPoint presentations and course credit. In addition to building upon
material from traditional coursework and staying current on the
literature, Journal Clubs help strengthen skills in critical thinking,
communication, and integrating the literature {[}11{]}. Since most
Journal Clubs occur on a regular schedule, they are effective by virtue
of repeated practice. With this model in mind, over the past four years
we have experimented with creating a Code Club model with the goal of
improving reproducible data analysis skills in a laboratory environment.

@@

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}12{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using R syntax from either the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. We generally
find that preparing a Code Club session takes similar effort to
preparing a Journal Club presentation. Our Code Clubs typically have 7
to 10 participants, but the inherent ``think-pair-share'' approach
should allow it to be scaled to groups of variable sizes {[}13{]}.

We continue to experiment with approaches for running Code Club. For
example, during the COVID-19 pandemic in 2020 when many research labs
were shuttered, PDS started posting Code Club materials
(\url{https://www.riffomonas.org/code_club/}) that individuals or
research groups can use to practice their coding skills. During this
time, we continued our group's Code Club sessions using the breakout
room and screen sharing features in Zoom. Regardless of our approach to
leading a Code Club, we have learned that it is critical for the
presenter to clearly articulate their goals and facilitate participant
engagement. Although some Code Club sessions may be more experimental
than others, on the whole they are a critical tool to train bench
scientists in reproducible data analysis practices. We have provided
some examples of successful topics in Table 1. We have summarized the
results of our own experimentation as Code Club presenters and
participants into Ten Rules. The first 3 rules apply to all
participants, Rules 4 through 8 are targeted to presenters, and the last
two for non-presenters. As the rules describe, this model can easily be
adapted to groups of non-biologists or those with stronger backgrounds
in computer science.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. Beginners will benefit from sessions that cover fundamental
concepts and functions. The benefits for more experienced participants
include the improved understanding of concepts by teaching and breaking
down problems into simpler elements. Furthermore, they can benefit by
contributing to an environment of learning that they previously
benefited from. In addition, instead of focusing on core functions, the
group could balance Code Club sessions that cover basics with those that
introduce new methods and packages to the group. We have also identified
several strategies to overcome the challenges presented by participants
at varying skill levels. Central to the Code Club format is the use of
paired-programming {[}15{]}. Instead of letting participants form their
own pairs, the presenter can select pairs of participants with either
similar or differing skill levels, depending on the presenter's goals.
Partnerships between those with similar skill levels requires the
presenter to design appropriate activities for each skill level. We have
found that commenting code is a good skill for beginners since it forces
them to dissect and understand a code chunk line-by-line. It also
reinforces the value of commenting as they develop their skills and
independence. An advantage of forming partnership between people with
disparate skill levels is that it is more likely for groups to provide
the presenter with a diverse range of methods that achieve the same
result. This approach to pairing also helps to graft new members that
have emerging programming skills. Regardless of how partners are
selected, consider asking the pairs to identify a navigator and a driver
{[}16{]}. The driver types at the computer while the navigator tells
them what to type, thus ensuring participation of both partners. Midway
through the activity, the presenter can have the partners switch roles.
Intentionally forming pairs can also engineer group interactions by
avoiding potentially disruptive partnerships or pairing reliable role
models with new group members.
